---
title: "ggplot"
author: "Qimin Zhang"
date: "9/26/2019"
output: github_document
---
# Lecture 1
```{r setup, include=FALSE, message=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
library(ggridges)
```

```{r, collapse = TRUE, message = FALSE, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```
Create a ggplot.
```{r, message = FALSE, warning = FALSE}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```
Alternative way to make this plot.
```{r, message = FALSE, warning = FALSE}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
    geom_point()
```
Saving initial plots, mostly I don't do this.
```{r, message = FALSE, warning = FALSE}
scatterplot = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
    geom_point()

scatterplot
```
Adding color...
```{r, message = FALSE, warning = FALSE}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
    geom_point(aes(color = name), alpha = .4)
```
Why do 'aes' positions matter? If I use 'color =' in a global scale, then all plots in this graph would be influenced.
```{r, message = FALSE, warning = FALSE}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
    geom_point(alpha = .4) +
    geom_smooth(se = FALSE)
```
Facet!
```{r, message = FALSE, warning = FALSE}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
    geom_point(alpha = .4) +
    geom_smooth(se = FALSE) +
    facet_grid(. ~ name)
```
Let's make some interesting plots.
```{r, message = FALSE, warning = FALSE}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE)

ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```
Only showing the smooth curve?
```{r, message = FALSE, warning = FALSE}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
```
```{r, message = FALSE, warning = FALSE}
ggplot(weather_df, aes(x = date, y = tmax)) + 
  geom_hex()
```
More kinds of plots.
```{r, message = FALSE, warning = FALSE}
#Histogram
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram()

ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```
```{r, message = FALSE, warning = FALSE}
#Density plot
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```
```{r, message = FALSE, warning = FALSE}
#Boxplot
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_boxplot()
```
```{r, message = FALSE, warning = FALSE}
#Violin plot
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 4)
```
```{r, message = FALSE, warning = FALSE}
#Ridge plot
ridge_temp =
  ggplot(weather_df, aes(x = tmax, y = name)) + 
    geom_density_ridges(scale = .85)
ggsave("gglpot_tmp_riage.pdf", ridge_temp)
```
# Lecture 2
Labels and ticks.
```{r,message=FALSE,warning=FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp",
    y = "Maximum Temp",
    caption = "Data from the rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15", "-5", "20")
  )
```
```{r,message=FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp",
    y = "Maximum Temp",
    caption = "Data from the rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15", "-5", "20")
  ) +
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```
or scale_y_sqrt().

Colors and themes.
```{r,warning=FALSE}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(name = "Location", h = c(100, 300))
```

```{r,warning=FALSE}
gg_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

gg_temp_plot
```
```{r, warning=FALSE}
gg_temp_plot + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```
Don't reverse the order!

More than one dataset.
```{r,message=FALSE}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```
To assign plot of sigle dataset color.
```{r}
waikiki %>%
  ggplot(aes(x = date, y = tmax)) +
  geom_point(color = "red")
```
```{r, warning=FALSE}
library(patchwork)

tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```
Data manipulation.
```{r,warning=FALSE}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
```{r,warning=FALSE}
weather_df %>%
  mutate(
    name = factor(name),
    name = reorder(name, tmin)
    ) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
```{r,warning=FALSE}
weather_df %>%
  select(name, tmax, tmin) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```
```{r}
pup_data = 
  read_csv("../data_wrangling_1/data_wrangling_1/data_import_examples//FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("../data_wrangling_1/data_wrangling_1/data_import_examples//FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, day_of_tx, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```
# Lecture 3
Create the weather data by month.
```{r, message=FALSE, warning=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, date, month, everything())
```
Group by yand counting.
```{r}
weather_df %>%
  group_by(name, month)
```
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(n_obs = n(),
            n_unique = n_distinct(date))
```
```{r}
weather_df %>%
  count(name, month) %>%
  knitr::kable()
```
2*2 tables
```{r}
weather_df %>%
  mutate(
    cold = case_when(
      tmax < 5 ~"cold",
      tmin > 25 ~ "hot",
      TRUE ~ ""
    )
  ) %>%
  group_by(name, cold) %>%
  count() %>%
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```
General summaries.
```{r}
weather_df %>%
  drop_na() %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax)
  ) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```
```{r}
weather_df %>%
  drop_na() %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)
```

```{r}
weather_df %>%
  drop_na() %>%
  group_by(name) %>%
  mutate(centered_tmax = tmax - mean(tmax)) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax)) %>%
  filter(temp_ranking == 1)
```
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax)) %>%
  summarize(temp_change_sd = sd(temp_change, na.rm = TRUE),
            temp_change_max = max(temp_change, na.rm = TRUE))
```


